Cluster Detection from three-dimensional Lensing Mass Map

### 2021-01-09 doc/paper_ms_method_HSCY1/ms.tex
update summary

\begin{abstract}
A new method is developed to reconstruct three-dimensional mass maps from
photometric weak-lensing shear measurements. The three-dimensional mass map is
modeled as a summation of the NFW basis atoms, which have two-dimensional
multi-scale NFW surface density profiles on the transverse plane and
one-dimensional Dirac delta functions in the line of sight direction.  The
adaptive lasso algorithm is applied to find a sparse reconstruction of the mass
maps.
We study the performance of three-dimensional cluster detection from the
reconstructed mass map with simulations that applies shear distortions from
isolated halo to HSC-like galaxy shapes with realistic photometric redshift
uncertainties .
Our findings are summarized as follows:
1) The lasso reconstructed suffers from a smear of structure in the line of
sight direction even in the absence of shape noise, and in contrast, the
adaptive lasso algorithm efficiently removes the line of sight smear.
2) The algorithm is able to detect halo with minimal mass limits of $10^{14.0}
M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low
($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts,
respectively, with a false detection of 0.022/deg$^2$.
3) The estimated redshift of the halos detected from the reconstructed mass
maps are lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$). The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$.
\end{abstract}

@@

We develop a novel method to reconstruct $3$-D density contrast maps from
weak-lensing shear measurements and photometric redshift estimations.  Our
method models $3$-D density contrast maps as summations of NFW atoms with
different comoving radii.  With the prior assumption that the clumpy masses
sparsely distribute in the $3$-D space, the density map is reconstructed
using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}. The method is
tested with realistic simulations using HSC-like shape estimation error and
photo-$z$ uncertainty.

The findings of this paper are summarized as follows:

(i) The lasso algorithm's solution suffers from a smear of structure in the
line of sight direction even in the absence of shape noise, and the adaptive
lasso algorithm efficiently removes the line of sight smear of structure.

(ii) The algorithm is able to detect halo with minimal mass limits of
$10^{14.0} M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for
the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$)
redshifts, respectively, with a false detection of 0.022/deg$^2$.

(iii)The estimated redshifts of the halos detected from the reconstructed mass
maps are lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$).  The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$.

We will apply the method to the shear measurements of the HSC survey
\citep{HSC1-catalog,FPFSHSC1-Li2020} to perform galaxy cluster detection in our
future work.

>>>>

\begin{abstract}
A new method is developed to reconstruct high-resoluion three-dimensional mass
maps from photometric weak-lensing shear measurements. The three-dimensional
mass map is modeled as a summation of the NFW basis atoms, which have
two-dimensional multi-scale NFW surface density profiles on the transverse
plane and one-dimensional Dirac delta functions in the line of sight direction.
The adaptive lasso algorithm is applied to find a sparse reconstruction of the
mass maps.
We study the performance of three-dimensional cluster detection from the
reconstructed mass map with simulations that applies shear distortions from
isolated halo to HSC-like galaxy shapes with realistic photometric redshift
uncertainties .
Our findings are summarized as follows:
1) The lasso reconstructed suffers from a smear of structure in the line of
sight direction even in the absence of shape noise, and in contrast, the
adaptive lasso algorithm efficiently removes the line of sight smear.
2) The algorithm is able to detect halo with minimal mass limits of $10^{14.0}
M_{\odot}/h$, $10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low
($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts,
respectively, with an average false detection of 0.022/deg$^2$.
3) The estimated redshift of the halos detected from the reconstructed mass
maps are lower than the true redshift by about $0.03$ for halos at low
redshifts ($z\leq 0.4$). The relative redshift bias is below $0.5\%$ for halos
at $0.4<z\leq 0.85$. The standard deviation of the redshift estimation is $0.092$.
\end{abstract}

@@

We have developed a novel method to reconstruct high-resolution $3$-D density
contrast maps from weak-lensing shear measurements and photometric redshift
estimations.  Our method models $3$-D density contrast maps as summations of
NFW atoms with different comoving radii.  With the prior assumption that the
clumpy masses sparsely distribute in the $3$-D space, the density map is
reconstructed using the adaptive lasso algorithm \citep{AdaLASSO-Zou2006}.
We find that the lasso algorithm's solution suffers from a smear of structure
in the line of sight direction even in the absence of galaxy shape noise, and
the adaptive lasso algorithm efficiently removes the line of sight smear of
structure.

We study the performance of cluster detection from the reconstructed $3$-D mass
maps using simulations that apply shear distortions from isolated NFW halos to
HSC-like shapes and photo-$z$ uncertainties.  We conclude that the algorithm is
able to detect halo with minimal mass limits of $10^{14.0} M_{\odot}/h$,
$10^{14.7} M_{\odot}/h$, $10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median
($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.85$) redshifts, respectively, with an
average false detection of 0.022 deg$^{-2}$. The estimated redshifts of the halos
detected from the reconstructed mass maps are lower than the true redshift by
about $0.03$ for halos at low redshifts ($z\leq 0.4$).  The relative redshift
bias is below $0.5\%$ for halos at $0.4<z\leq 0.85$. The standard deviation of
the redshift estimation is $0.092$.

We will apply the $3$-D mass map reconstruction to the shear measurements of
the HSC survey \citep[e.g.,][]{HSC1-catalog,FPFSHSC1-Li2020} to perform galaxy
cluster detection in our future work.
